---
title: "German Credit Analysis"
author: "Conor Ryan - 16340116"
date: "22/12/2021"
output: pdf_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Clearing environment}
# Ensuring environment is clear
rm(list = ls())
```

```{r Loading required packages}
# Loading all required packages
pacman::p_load(tidyverse,scales,cowplot,DataExplorer,kableExtra,
               ROCR,rpart,yardstick)
```

```{r Set Current Working Directory}
# Setting the current working directory
setwd("C:\\Users\\conor\\OneDrive\\Documents\\Data Analytics Masters\\Stage 1\\Semester 1\\Data Prog with R - STAT40730\\Assignments\\R Project")
```

```{r Setting Random Seed}
# Setting the random seed to ensure results are reproducible
set.seed(1759)
```

# Part 1: Analysis
## Overview of the data set

The data-set I have chosen to perform my analyses on is the German Credit Data data-set which is available from the UCI (University of California Irvine) machine learning data repository via the following link   
[https://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data](https://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data).

The data-set was compiled by Professor Hans Hofmann from the institute of Statistics and Econometrics at the University of Hamburg and contains information on 1,000 German borrowers collected between 1973 and 1975. The data set classifies borrowers as having "good" or "bad" credit ratings and also provides 20 additional variables for each credit customer e.g. duration of loan, customer's credit history, purpose and amount of the loan etc. Further information regarding this data set can be found via this link      
[https://archive.ics.uci.edu/ml/datasets/Statlog+%28German+Credit+Data%29](https://archive.ics.uci.edu/ml/datasets/Statlog+%28German+Credit+Data%29).

My analysis will consist of a portfolio overview which will attempt to gain a better understanding of the characteristics of the 1,000 customers in the data set. Firstly I will read in the German Credit data set. The data is saved as a ".data" file type which is simply a text file with the values separated by a certain character, a space in this case. I read the data directly from the UCI data repository website so that my analysis is more easily reproducible.

*Please note during this project I have used several packages for various reasons, (formatting tables, axes on graphs, plotting graphs etc.) that were not used extensively throughout the module. I have included a bibliography in the appendices at the end of my report where I cite these packages and explain the rationale for using them*  

```{r Reading in German Credit data}
# Reading in the data from the UCI ML data repository
german_df <- 
  read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data",
             sep = " ")

# Displaying the first 5 rows of the data set
# I use the kableExtra package to "scale-down" the table so it fits within the PDF
kbl(head(german_df,5),
    caption = "German Credit (Raw) Data",
    position = "h") %>% 
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r Size of German Credit data}
# Displaying size of the data set
print(c("Rows" = nrow(german_df),"Columns" = ncol(german_df)))
```

As we can see from the output above there are **`r nrow(german_df)`** rows and **`r ncol(german_df)`** columns (20 customer characteristic columns and 1 binary credit rating) in the German Credit data set. 

## Data pre-processing
### Adding column headers and coverting to factors

We notice that many of the categorical variables have been coded in a way which makes it impossible to determine their meaning without reading the data set description document I have mentioned previously. There are also no column headers present which makes it even more difficult to analyze the data. Therefore, before performing any analysis I will first rename the column headings and values to make the data set more usable.

```{r Renaming German Credit columns}
# Creating a list of column names (obtained from description document mentioned in introduction)
german_colnames <- c("Checking_Acc_Bal","Loan_Duration","Credit_History",
                     "Loan_Purpose","Credit_Amount","Savings_Acc_Bal",
                     "Employment_Duration","Payments_Percent_Of_Income",
                     "Personal_Status","Other_Debtors","Resident_Since",
                     "Property","Age","Other_Install_Plans",
                     "Housing","Num_Existing_Credits","Employment_Status",
                     "Num_Dependents","Telephone","Foreign_Worker",
                     "Credit_Rating")

colnames(german_df) <- german_colnames
```

```{r Renaming German Credit Values}
german_df <- german_df %>%
  # Renaming each of the coded categorical variables in the data set
  # To improve readability and user-friendliness of the data
  mutate(Checking_Acc_Bal = factor(Checking_Acc_Bal,
                                   levels = c("A14","A11","A12","A13"),
                                   labels = c("No Checking Account",
                                              "< 0 DM",
                                              "< 200 DM",
                                              ">= 200 DM"),
                                   ordered = TRUE),
         Credit_History = factor(Credit_History,
                                 levels = c("A30","A31","A32","A33","A34"),
                                 labels = c("All credits paid duly",
                                            "All credits at this bank paid duly",
                                            "Existing credits paid duly until now",
                                            "Delayed payments in the past",
                                            "Critical account"),
                                 ordered = FALSE),
          Loan_Purpose = factor(Loan_Purpose,
                                levels = c("A40","A41","A42","A43","A44",
                                            "A45","A46","A47","A48","A49","A410"),
                                labels = c("Car (new)","Car (used)","Furniture/Equipment",
                                            "Radio/TV","Domestic Appliance","Repairs",
                                            "Education","Holiday","Retraining",
                                            "Business","Other"),
                                ordered = FALSE),
         Savings_Acc_Bal = factor(Savings_Acc_Bal,
                                 levels = c("A65","A61","A62","A63","A64"),
                                 labels = c("No Savings Account",
                                            "< 100 DM",
                                            "100 - 500 DM",
                                            "500 - 1000 DM",
                                            "> 1000 DM"),
                                 ordered = TRUE),
         Employment_Duration = factor(Employment_Duration,
                                      levels = c("A71","A72","A73","A74","A75"),
                                      labels = c("Unemployed",
                                                 "< 1 year",
                                                 "< 4 years",
                                                 "< 7 years",
                                                 ">= 7 years"),
                                      ordered = TRUE),
         Personal_Status = factor(Personal_Status,
                                 levels = c("A91","A92","A93","A94","A95"),
                                 labels = c("Male - Divorced/Separated",
                                            "Female - Divorced/Separated/Married",
                                            "Male - Single",
                                            "Male - Married/Widowed",
                                            "Female - Single"),
                                 ordered = FALSE),
         Other_Debtors = factor(Other_Debtors,
                                levels = c("A101","A102","A103"),
                                labels = c("No Other Debtors",
                                           "Co-Applicant",
                                            "Guarantor"),
                                ordered = FALSE),
         Property = factor(Property,
                           levels = c("A121","A122","A123","A124"),
                           labels = c("Real Estate",
                                      "Savings Agreement/Life Insurance",
                                      "Car/Other",
                                      "Unknown/No Property"),
                           ordered = FALSE),
         Other_Install_Plans = factor(Other_Install_Plans,
                                      levels = c("A141","A142","A143"),
                                      labels = c("Bank",
                                                 "Stores",
                                                 "None"),
                                      ordered = FALSE),
         Housing = factor(Housing,
                          levels = c("A151","A152","A153"),
                          labels = c("Rent",
                                     "Owned",
                                     "Free"),
                          ordered = FALSE),
         Employment_Status = factor(Employment_Status,
                                    levels = c("A171","A172","A173","A174"),
                                    labels = c("Unemployed/Unskilled - Non-Resident",
                                               "Unskilled - Resident",
                                               "Skilled Employee/Official",
                                               "Management/Self-Employed/Highly-Skilled/Officer"),
                                    ordered = FALSE),
         Telephone = factor(Telephone,
                            levels = c("A191","A192"),
                            labels = c("N","Y"),
                            ordered = FALSE),
         Foreign_Worker = factor(Foreign_Worker,
                                 levels = c("A201","A202"),
                                 labels = c("Y","N"),
                                 ordered = FALSE),
         Credit_Rating = factor(Credit_Rating,
                                levels = c(1,2),
                                labels = c("Good","Bad"),
                                ordered = FALSE))

# Displaying the first 5 rows and 5 columns of the dataset
# I use the kableExtra package to produce a nicely formatted output
kbl(head(german_df[,1:5],5),
    caption = "German Credit (Processed) Data",
    position = "h") %>% 
  kable_styling(latex_options = c("striped","scale_down"))
```

*Just to point out above that the "DM" characters within the checking account balance stands for Deutsche-Mark or D-Mark which was the currency used in Germany before the arrival of the euro in 2002.* 

### Removing unnecessary variables

Now that we have converted the previously coded categorical variables to user-friendly factors the next step is to remove the unnecessary and unusable variables for which we have no use for or where the descriptions of the variables in the data dictionary are unclear. The variables I have chosen to remove are the following;  

1. **Property** - Description of variable unclear  
2. **Other Installment Plans** - Description of variable unclear  
3. **Number of Existing Credits** - Description of variable unclear  
4. **Telephone** - Unnecessary  
5. **Other Debtors** - Description of variable unclear  

```{r Removing variables from German Credit data}
# Removing un-clear / unnecessary variables from the data
german_df <- german_df %>% 
  select(-Property,-Other_Install_Plans,-Num_Existing_Credits,
         -Telephone,-Other_Debtors)

# Printing the updated dimensions of the data set 
print(c("Rows" = nrow(german_df),"Columns" = ncol(german_df)))
```

### Deriving additional variables

Next I would like to derive a few additional columns based on some of the variables within the data set. Firstly, I wish to split out the "Personal_Status" variable to produce a gender column. For the purposes of this analysis I assume there are no widowed females in the data set as this cohort is not included in any of the 5 values that the personal status variable can take.   

I will also produce a banded age variable derived from the age column as I believe this will be useful in producing plots and graphical summaries. I produce a descriptive summary of the age variable to show the minimum and maximum values in order to inform my choice of age banding. As we can see below the minimum age of borrower in the data set is 19 and the maximum is 75.  

```{r Summary of Age variable German Credit data}
# Firstly I produce a summary of the age variable to inform my age banding column
# (i.e. min and max ages within the data)

print(summary(german_df$Age))
```

```{r Derived variables German Credit data}
# Producing gender and age banding variables based on the personal status and
# age variables respectively
german_df <- german_df %>% 
  mutate(Gender = case_when(grepl(pattern = "Female",
                                  x = Personal_Status,ignore.case = TRUE) ~ "Female",
                             grepl(pattern = "Male",
                                   x = Personal_Status,ignore.case = TRUE) ~ "Male",
                             TRUE ~ "Unknown"),
         Gender = factor(Gender),
         Age_Banding = case_when(# In case of any missing values set the age band to "unknown age"
                                 is.na(Age) ~ "Unknown Age",
                                 Age <= 20 ~ "0 - 20 years",
                                 Age <= 30 ~ "21 - 30 years",
                                 Age <= 40 ~ "31 - 40 years",
                                 Age <= 50 ~ "41 - 50 years",
                                 Age <= 60 ~ "51 - 60 years",
                                 Age <= 70 ~ "61 - 70 years",
                                 Age <= 80 ~ "71 - 80 years",
                                 Age > 80 ~ "80+ years"),
         Age_Banding = factor(Age_Banding,
                              levels = c("Unknown Age",
                                         "0 - 20 years",
                                         "21 - 30 years",
                                         "31 - 40 years",
                                         "41 - 50 years",
                                         "51 - 60 years",
                                         "61 - 70 years",
                                         "71 - 80 years",
                                         "80+ years"),
                              ordered = TRUE))
```

### Checking for missing data

The final task I will perform before going on and analyzing the data set is to check for any missing values. Missing values can cause certain functions (such as means, totals, averages etc.) to not work and obscure summaries and graphs. If we have missing values in the dataset we need to figure out a way of dealing with them such as removing any observation which contains missing data or imputing the missing data points using a variety of methods (e.g. mean / null imputation).  
As we can see below we have no missing data points within the data set and therefore we are ready to move onto performing our analyses.

```{r Check for missing values German Credit data}
# Calculating the total number of missing values in the data
no_missing_vals <- sum(is.na(german_df))

# Writing an if statement to tell us whether there is missing data or not
if (no_missing_vals == 0) {
  print("There are no missing values in the German Credit data set")
} else {
  print(paste0("There are ",no_missing_vals," missing values in the German Credit data set"))
}
```

## Analysis

Since we have completed the data pre-processing stage, we are ready to move onto the analysis of the data set. The variable of interest is whether or not the customer has a "good" or "bad" credit rating and if we were to perform a classification exercise using machine learning on this dataset then the "Credit Rating" variable would be our dependent or target variable. This means that we would be interested in predicting / explaining a customer's credit rating based on the remaining variables in the data (explanatory variables). Therefore, I will first calculate the proportion of our customers who have a "good" credit rating versus those who don't.

```{r Summary of Credit Rating German Credit data}
# Producing simple summary of the credit rating variable
print(summary(german_df$Credit_Rating) / length(german_df$Credit_Rating))

```

As we can see above `r sum(german_df$Credit_Rating == "Good")/length(german_df$Credit_Rating) * 100`% of customers have good credit ratings whereas `r (1 - sum(german_df$Credit_Rating == "Good")/length(german_df$Credit_Rating)) * 100`% have bad credit ratings. This shows our data set is slightly unbalanced with respect to credit rating.  

Now I will display the structure of the data set to determine what kind of variables we have e.g. numerical, nominal, ordinal etc. Then I will present a summary of each of these variables depending on their data type in order to explore how these variables are distributed. 

```{r Displaying the structure of the dataset}
# Displaying the structure of the data set
str(german_df)
```
```{r Summaries of nominal variables}
# Producing summary proportions for nominal (non-ordered) variables

# Iterating over the columns in the data set (after dropping the credit rating)
for (col in colnames(select(german_df,-Credit_Rating))) {
  # If the variable is a factor then we print the summary
  # (dividing by total rows to give proportions)
  if (class(german_df[,col])[1] == "factor") {
    cat(paste0("**",col,"**"),"\n")
    print(round(summary(german_df[,col]) / nrow(german_df),3))
    cat("\n")
  } else {
    # else do nothing
  }
}
```
```{r Graphical summaries of nominal variables, fig.height=10,fig.width=15,fig.align="center"}
# Producing bar charts for nominal variables

# Credit history
cred_hist_bar <- ggplot(data = german_df,aes(x = Credit_History,fill=Credit_Rating)) +
  geom_bar(position = "stack",aes(y = ..count..),col = "black") +
  # adding counts of "good" & "bad" credit ratings
  geom_text(aes(label = ..count..),stat = "count",position = position_stack(vjust = 0.5)) +
  # adding a title
  ggtitle("Credit History") +
  xlab("Credit History") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14),
        axis.text.x = element_text(angle = 30,vjust = 0.7),
        legend.position = c(.15,.85)) +
  # manually adjusting x value labels for aesthetics
  scale_x_discrete(labels = c("Creds paid duly","Creds bank paid duly",
                              "Ex. creds paid duly","Delayed payments",
                              "Critical account")) +
  # adjusting colours of "good" and "bad" credit rating
  scale_fill_manual(values = c("green3","red3"))

# Housing
house_bar <- ggplot(data = german_df,aes(x = Housing,fill=Credit_Rating)) +
  geom_bar(position = "stack",aes(y = ..count..),col = "black") +
  # adding counts of "good" & "bad" credit ratings
  geom_text(aes(label = ..count..),stat = "count",position = position_stack(vjust = 0.5)) +
  # adding a title
  ggtitle("Housing Status") +
  xlab("Housing Status") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14),
        legend.position = c(.15,.85)) +
  # adjusting colours of "good" and "bad" credit rating
  scale_fill_manual(values = c("green3","red3"))

# Employment status
employ_stat_bar <- ggplot(data = german_df,aes(x = Employment_Status,fill=Credit_Rating)) +
  geom_bar(position = "stack",aes(y = ..count..),col = "black") +
  # adding counts of "good" & "bad" credit ratings
  geom_text(aes(label = ..count..),stat = "count",position = position_stack(vjust = 0.5)) +
  # adding a title
  ggtitle("Employment Status") +
  xlab("Employment Status") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14),
        legend.position = c(.15,.85)) +
  # manually adjusting x value labels for aesthetics
  scale_x_discrete(labels = c("Unemployed","Unskilled",
                              "Skilled","Highly-Skilled")) +
  # adjusting colours of "good" and "bad" credit rating
  scale_fill_manual(values = c("green3","red3"))

# Gender
gender_bar <- ggplot(data = german_df,aes(x = Gender,fill=Credit_Rating)) +
  geom_bar(position = "stack",col = "black") +
  # adding counts of "good" & "bad" credit ratings
  geom_text(aes(label = ..count..),stat = "count",position = position_stack(vjust = 0.5)) +
  # adding a title
  ggtitle("Gender") +
  xlab("Gender") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14),
        legend.position = c(.15,.85)) +
  # adjusting colours of "good" and "bad" credit rating
  scale_fill_manual(values = c("green3","red3"))

# plotting all charts in the one figure using plot grid
plot_grid(cred_hist_bar,house_bar,employ_stat_bar,gender_bar,
          nrow = 2,ncol = 2)

```

Above I have produced the proportion summaries for each of the nominal variables in the data set. I have also created stacked bar charts for the credit history, housing, employment status and gender variables which I have split by credit rating. It is important to analyse the breakdown of each of these variables in terms of the customer's credit rating as we are interested in identifying patterns in the data which may influence a given customer's credit score. 

Observing the numerical summary output we can see that over half of the customers have either all their credits paid on time (at the current or different bank) or their existing credits paid on time up to the point of data collection (`r round((sum(german_df$Credit_History %in% c("All credits paid duly","All credits at this bank paid duly","Existing credits paid duly until now")) / nrow(german_df)) * 100,0)`%). There are however, `r round((sum(german_df$Credit_History == "Critical account") / nrow(german_df)) * 100,0)`% of customers whose account histories are critical. There is not much information given on the critical account cohort in the data dictionary and therefore it is difficult to understand what is meant by "critical". Observing the bar chart for the credit history we can see that the majority of the "critical account" cohort have "good" credit ratings (243 vs 50). We note that there are a large number of "good" credit ratings in the "Existing credits paid duly til now" cohort and surprisingly there is a large proportion of "bad" credit ratings in the two "paid duly" cohorts both at this bank and at another bank.      

We can see that the majority of loans are being used either to purchase a radio or TV (`r round(sum(german_df$Loan_Purpose == "Radio/TV")/nrow(german_df) * 100,0)`%) or to buy a new car (`r round(sum(german_df$Loan_Purpose == "Car (new)")/nrow(german_df) * 100,0)`%). On the other hand there are not many loans being used to go on holidays, purchase domestic appliances or for retraining purposes (all <= 1%). I will dive a bit deeper into the purposes of the loans further on in my analysis.  

In terms of gender demographics, `r round(sum(german_df$Gender == "Male")/nrow(german_df) * 100,0)`% of the portfolio are males and of those males `r round(sum(german_df$Personal_Status == "Male - Single")/sum(german_df$Gender == "Male") * 100,0)`% are single. We have no single females in the data set whatsoever. Potentially this could be due to a lack of women's civil rights in Germany during the period 1973-1975 as it was not until 1977 that women were allowed to work without first getting permission from their husbands (Wikipedia - [https://en.wikipedia.org/wiki/Timeline_of_women%27s_legal_rights_(other_than_voting)_in_the_20th_century]). Again, there is no obvious pattern with regards to credit ratings and gender as you would expect intuitively.

We see that most of the customer's in this portfolio are owners of their place of residence and we also see that home ownership has a far greater proportion of "good" credit ratings compared with renting or living for free (assuming in their parent's house). If you own your own house it is likely you have either successfully paid off a mortgage or are currently paying one off and therefore you would have a better credit rating than people who do not own their own house. 

One of the proportions that really stands out in the data set is the proportion of customers who are foreign workers. Foreign workers make up a massive `r round(sum(german_df$Foreign_Worker == "Y")/nrow(german_df) * 100,0)`% of the portfolio and this is something we must factor in when making any inferences about the German credit landscape as a whole. This is because our data is heavily biased towards foreign workers and may not represent a typical German credit portfolio. 

```{r Summaries of ordinal variables}
# Producing summary proportions for ordinal variables

# Iterating over the columns in the data set (after dropping the credit rating)
for (col in colnames(select(german_df,-Credit_Rating))) {
  # If the variable is an ordered factor then we print the summary 
  # (dividing by total rows to give proportions)
  if (class(german_df[,col])[1] == "ordered") {
    cat(paste0("**",col,"**"),"\n")
    print(round(summary(german_df[,col]) / nrow(german_df),3))
    cat("\n")
  } else {
    # else do nothing
  }
}
```

```{r Graphical summary of Ordinal variables, fig.height=12,fig.width=15,fig.align="center"}
# Plotting bar charts & densities of Checking account balance, Savings account balance,
# Duration of employment and Age banding

# Checking account balance
check_acc_bar <- ggplot(data = german_df,aes(x = Checking_Acc_Bal)) +
  geom_bar(aes(y = ..count..),fill = "yellow2",col = "black") +
  # adding a title
  ggtitle("Checking Account Balance (DM)") +
  xlab("Checking Account Balance") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14))

# Savings account balance
Savings_acc_bar <- ggplot(data = german_df,aes(x = Savings_Acc_Bal)) +
  geom_bar(aes(y = ..count..),fill = "steelblue2",col = "black") +
  # adding a title
  ggtitle("Savings Account Balance (DM)") +
  xlab("Savings Account Balance") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text.x = element_text(angle = 30,vjust = 0.85),
        axis.text = element_text(size = 14))

# Employment duration
employ_dur_bar <- ggplot(data = german_df,aes(x = Employment_Duration)) +
  geom_bar(aes(y = ..count..),fill = "darkmagenta",col = "black") +
  # adding a title
  ggtitle("Employment Duration") +
  xlab("Employment_Duration") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 14))

# Age banding
age_band_bar <- ggplot(data = german_df,aes(x = Age_Banding)) +
  geom_bar(aes(y = ..count..),fill = "coral",col = "black") +
  # adding a title
  ggtitle("Age Banding") +
  xlab("Age Band") +
  ylab("No. Customers") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text.x = element_text(angle = 30,vjust = 0.85),
        axis.text = element_text(size = 14))


# plotting all of the bar charts within the same plot using plot_grid function
plot_grid(check_acc_bar,employ_dur_bar,Savings_acc_bar,age_band_bar,
          nrow = 2,
          ncol = 2)

```

I have produced similar proportion summaries for the ordinal variables along with graphical summaries in the form of bar charts as shown above. We can see that c.39% of customers do not have a checking account and a further c.27% are overdrawn on their checking account (balance < 0 dms). Furthermore, c.79% of customers either do not have a savings account or have less than 100 D-Marks saved. This could cause alarm to a portfolio manager especially if the loan happens to be unsecured as if the loan were to default the bank would have nothing to repossess or sell in order to make back some of its loss.  

A more positive statistic in terms of credit risk is that c.77% of customers have been in their current employment for over 4 years which bodes well in terms of job security and for the bank manager's point of view, this corresponds to a steadier stream of income which can be used to pay back the debts owed to the bank. We can see in terms of age demographics that the vast majority of customer's are between the ages of 21 and 50 (c.87%).


```{r Summaries of numerical variables}
# Producing summary proportions for numerical variables

# Iterating over the columns in the data set (after dropping the credit rating)
for (col in colnames(select(german_df,-Credit_Rating))) {
  # If the variable is numerical then we print the summary 
  if (!class(german_df[,col])[1] %in% c("factor","ordered")) {
    cat(paste0("**",col,"**"),"\n")
    print(c(summary(german_df[,col]),
            "Variance" = var(german_df[,col]),
            "St.Dev" = sd(german_df[,col]),
            "IQR" = IQR(german_df[,col])))
    cat("\n")
  } else {
    # else do nothing
  }
}
```

```{r graphical summaries of numerical variables,fig.height=15,fig.width=15,fig.align="center"}
# Plotting histograms / bar charts of Loan Duration,Payments as a % of income,
# Age and Resident since variables

# First plot
# Simple histogram of loan duration

loan_duration_hist <- ggplot(data = german_df,aes(x = Loan_Duration)) +
  geom_histogram(aes(y = ..count..),bins = 12,fill = "steelblue2",col = "black") +
  # adding mean line 
  geom_vline(aes(xintercept = mean(Loan_Duration),col="mean"),size=1.25,linetype="dashed") +
  # adding median line
  geom_vline(aes(xintercept = median(Loan_Duration),col="median"),size=1.25,linetype="dashed") +
  # setting the colours for the mean and median lines
  scale_color_manual(name="Descriptive Stats",values = c(mean = "red",median = "green")) +
  # adding a title
  ggtitle("Loan Duration (Months)") +
  ylab("No. Customers") +
  # formatting the x axis with "scale" package
  scale_x_continuous(name = "Loan Duration (Months)",
                     breaks = seq(0,72,4)) +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        legend.position = c(.8,.9),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

# Second plot
# Simple bar chart of Payments as a % of income

# Dividing payments percent of income by 100 to convert to  a %age value
pay_percent_income_bar <- ggplot(data = german_df,aes(x = Payments_Percent_Of_Income/100)) + 
  geom_bar(aes(y = ..count..),fill = "yellow2",col = "black") +
   # adding mean line 
  geom_vline(aes(xintercept = mean(Payments_Percent_Of_Income/100),col="mean"),size=1.25,linetype="dashed") +
  # adding median line
  geom_vline(aes(xintercept = median(Payments_Percent_Of_Income/100),col="median"),size=1.25,linetype="dashed") +
   # setting the colours for the mean and median lines
  scale_color_manual(name="Descriptive Stats",values = c(mean = "red",median = "green")) +
  # adding a title
  ggtitle("Repayments as a % Income") +
  ylab("No. Customers") +
  # formatting the x axis with "scale" package
  scale_x_continuous(labels =  percent,
                   name = "Repayments as a % Income") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        legend.position = c(.2,.9),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

# Third plot
# Simple histogram of Customer age

age_hist <- ggplot(data = german_df,aes(x = Age)) +
  geom_histogram(aes(y = ..count..),bins = 20,fill = "coral",col = "black") +
  # adding mean line 
  geom_vline(aes(xintercept = mean(Age),col="mean"),size=1.25,linetype="dashed") +
  # adding median line
  geom_vline(aes(xintercept = median(Age),col="median"),size=1.25,linetype="dashed") +
  # setting the colours for the mean and median lines
  scale_color_manual(name="Descriptive Stats",values = c(mean = "red",median = "green")) +
  # adding a title
  ggtitle("Customer Age") +
  ylab("No. Customers") +
  # formatting the x axis with "scale" package
  scale_x_continuous(name = "Age (Years)",
                     breaks = seq(10,80,5)) +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        legend.position = c(.8,.9),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

# Fourth plot
# Simple histogram of Resident since 
# (i.e. in current residence for x years)

resident_bar <- ggplot(data = german_df,aes(x = Resident_Since)) +
  geom_bar(aes(y = ..count..),fill = "darkmagenta",col = "black") +
  # adding mean line 
  geom_vline(aes(xintercept = mean(Resident_Since),col="mean"),size=1.25,linetype="dashed") +
  # adding median line
  geom_vline(aes(xintercept = median(Resident_Since),col="median"),size=1.25,linetype="dashed") +
  # setting the colours for the mean and median lines
  scale_color_manual(name="Descriptive Stats",values = c(mean = "red",median = "green")) +
  # adding a title
  ggtitle("Resident Since (Years)") +
  ylab("No. Customers") +
  xlab("Resident Since (Years)") +
  # adjusting titles and text size
  theme(plot.title = element_text(hjust = 0.5,size = 20),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        legend.position = c(.2,.9),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))


plot_grid(loan_duration_hist,pay_percent_income_bar,age_hist,resident_bar,
          nrow = 2, ncol = 2)
```

Finally, I have produced the summary statistics for the numerical variables in the German Credit data set. Starting with the duration of the loan we can see that the average term of a loan is 20.9 months. This is quite a short period of time particularly for loans, however if we observe the loan purpose variable we can see that none of these loans are mortgages which would generally have a term of at least 10 years. We also see that the max term is 72 months and therefore, we can conclude that we are looking at a short to medium term loan portfolio.  

I have also created graphical summaries (histograms and bar charts) of four of the numerical variables mentioned above, namely; Loan duration, payment as a % income, resident since (years) and customer age. While the numerical summaries provide a great amount of information, the histogram and bar charts shown above give a much clearer and illustrative view of the distribution of each of these variables. 

We can see from the plots above that both Age and Loan Duration are right-skewed. We observe nearly 300 loans in the 8-12 month range which is 30% of the entire portfolio. Interestingly in terms of age we can see a lot of the customers >250 are between the ages of 22 and 28. We can see the majority of customers spend 4% of their income on loan repayments, while this is the highest value within this data set, I believe 4% is still relatively low. Similarly over 400 customers have been in their current residence for 4 years.   

```{r Credit amount plots, fig.height=8,fig.width=15,fig.align="center"}
# First plot
# Simple histogram and density plot of the distribution of the credit amount
cred_amt_density <- ggplot(data = german_df,aes(x = Credit_Amount)) +
  geom_histogram(aes(y = ..density..),bins = 30,fill = "steelblue2") +
  geom_density(col = "firebrick1",size = 1.25) +
  ggtitle("Histogram & Density - Credit Amount") +
# Changing the y-axis to be in percentages (using scales package)
  scale_y_continuous(labels = percent,name = "Density") +
  scale_x_continuous(name = "Credit Amount (DM)",
                     breaks = seq(0,20000,2500)) +
  theme(plot.title = element_text(hjust = 0.5,size = 24),
        axis.title = element_text(size = 19),
        axis.text = element_text(size = 15))

# Second plot
# Cumulative distribution function of the credit amount variable
cred_amt_cdf <- ggplot(data = german_df,aes(x = Credit_Amount)) +
  stat_ecdf(geom = "step",pad = FALSE,col = "firebrick1",size = 1.25) +
  ggtitle("CDF - Credit Amount") +
# Changing the y-axis to be in percentage (using scales package) 
  scale_y_continuous(labels = percent,name = "CDF") +
  scale_x_continuous(name = "Credit Amount (DM)",
                     breaks = seq(0,20000,2500)) +
  theme(plot.title = element_text(hjust = 0.5,size=24),
        axis.title = element_text(size = 19),
        axis.text = element_text(size = 15))

# using plot grid from the cowplot package to arrange plots side by side
plot_grid(cred_amt_density,cred_amt_cdf)

```

Observing the statistics for the credit or loan amount we see the average loan amount is 3,271 dm whereas the min and max are 250 dm and 18,424 dm respectively. This indicates that the loans amounts are quite varied as we have a large range (c.18,150 dm) and standard deviation (2,822.74 dms). I have produced a density plot in order to visualise the distribution of the credit amount within the portfolio (shown above).

Looking at the histogram and density plot on the left we can see that the credit amount is heavily skewed to the right. This means that the majority of customer loans within the portfolio are for smaller amounts (<5000 dm). Viewing the cumulative distribution function (CDF) on the right we can see that just over half of all customer loans are for 2,500 dm or less, `r sum(german_df$Credit_Amount <= 2500)/nrow(german_df)*100`% to be precise.  

The distribution of loan amounts in the portfolio is consistent with the purposes for which the loans were taken out. As mentioned previously the majority of these loans were taken out to purchase either a radio or TV or to buy a new car. These are small - mid range purchases and is reflected in the skewed distribution of loan amounts. Had this portfolio be comprised of mortgages we would expect to observe a left-skewed distribution of loan amount. I will now further explore the make-up of this loan portfolio by looking more closely at the loan purpose variable. 

```{r Barplot loan purpose,fig.height=8,fig.width=15,fig.align="center"}
barplot_loan_purpose <- ggplot(data = german_df,aes(y = Credit_Amount,x = Loan_Purpose,fill=Loan_Purpose)) +
  geom_boxplot(col = "black") +
  ggtitle("Loan Purpose") +
# Formatting the y-axis using scales package
  scale_y_continuous(name = "Credit Amount (DM)",
                     breaks = seq(0,20000,2500)) +
  theme(plot.title = element_text(hjust = 0.5,size = 24),
        axis.title = element_text(size = 19),
        axis.text.x = element_text(angle = 30,vjust = 0.8),
        axis.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16))

barplot_loan_purpose
```

I have created a box-plot of the loan purpose variable against the credit amount of the loan, shown above. We can see that most of the cohorts are centred around the 2,500 dm level which is consistent with what we have seen in the previous density and CDF plots. Interestingly, the "other" cohort is by far the most varied with a very large inter-quartile range ranging from c.2,500 to 12,500 dm. Intuitively this makes sense as it's likely the "other" group is made up of many different loan purposes which would lead to increase variance in the loan amounts. We see that the "other" group also has the largest median value of c.7,000 dm.  

Another interesting insight gleaned from the boxplot above is the number of outliers we have within the "Radio/TV" group, our most popular group in the data set. We can see one of the outliers has a credit amount of over 15,000 dm! Immediately this would cause concern with a portfolio manager as it is unlikely even back in 1970s Germany that a radio or TV would cost more than a new car. This could be an error in the inputting of the data or the purpose of the loan being incorrectly classified as "Radio/TV".  

Comparing the "Car (new)" and "Car (used)" groups we can see another interesting anomaly. The median value for the used car group is much larger than that of a new car. This is not what we would expect to see in the data as cars generally depreciate in value after their original purchase and thus used cars tend to be much cheaper than new cars. Granted we do observe values in the new car group that are higher than the maximum used car amounts but in terms of the distributions we would expect to see the opposite of what we observe in this data set.

## Summary of Results

I have analyzed the German Credit data set using a variety of numerical and graphical summaries in order to get a better idea the characteristics within the portfolio of 1,000 customer loans and how these characteristics may affect a customer's credit rating. The 1,000 customer's are predominantly male (690) but the proportion of males vs females with "bad" credit ratings is roughly equal and therefore we can't say that gender is a contributing factor to credit score. The portfolio is a small-medium term loan portfolio and is made up predominantly of smaller loan amounts (skewed to the right). This is because the purpose of many of these loans are for items such as cars, radios & TVs and domestic appliances and not for homes or private jets. The average age of a customer is 35 and over 85% of customers are between the ages of 21 and 50. This means the majority of customers are in prime working age groups and should correspond to higher levels of repayment should they remain in employment. As at time of data collection 22 customers are unemployed and of these 7 have "bad" credit ratings.

We have also discovered that the duration of these loans is short with an average of 20.9 months and a maximum of 6 years. From a repayment ability perspective, most customers have been in their current employment for the last 4 years or more (c.77%) and on average have been resident in their current residence for 2.85 years. Both of these statistics demonstrate an ability of the majority of customer's to hold down a job and to pay their rent / mortgage repayments which lessens the risk of customers within the portfolio defaulting on their debt.

In the next part of the project I will demonstrate some of the functions within the "DataExplorer" package to show how much of the exploratory data analysis I have conducted here can be automated.

# Part 2: Functions/Programming
## Introduction
For the final part of my project I will write a function to perform three different classification algorithms and implement them on the German Credit data set. The main purpose of the data set and the reason it is available from UCI's machine learning repository, is to train and test various machine learning algorithms. The "Credit_Rating" variable is our target variable (this is the one we wish to explain / predict) and the remaining variables will be our explanatory variables (these are used to explain / predict the target variable). 

As the credit rating variable is binary and can take values of either "good" or "bad", we have a "Classification" problem. These are problems in which we wish to classify the target variable into certain cohorts based on the values of the explanatory variables. There are a plethora of classification algorithms available in various packages in R, however to keep it simple I have chosen to use just two and compare their results. The two algorithms I have chosen to implement are **logistic regression** and **decision tree**.  

I will create a function to perform each of the algorithms and record their respective results. Then I will create another function which runs each of the individual classification functions and collates the results of each one and producing an object of S3 class. Following this I will produce three distinct methods for this S3 class; a print method, a summary method and finally a plot method. Firstly, I must convert the credit rating to a binary variable in order to perform the two classification methods. We wish to predict whether or not a customer has a "bad" credit rating and therefore I map the value 1 to the "bad" cohort and 0 to the "good" cohort. 

```{r Converting credit rating to binary}
# converting the credit rating variable to binary
# We want to predict whether the customer has a bad credit rating
# therefore we set that to be 1 and good becomes 0
german_df <- german_df %>% 
  mutate(Credit_Rating = case_when(Credit_Rating == "Bad" ~ 1,
                                   Credit_Rating == "Good" ~ 0))
```

```{r Dropping the age banding column}
# I drop the age banding column as it is bound to be highly correlated with
# the age column (as its derived from this column)
german_df <- german_df %>% 
  select(-Age_Banding)
```

```{r Transform categorical variables into dummy variables}
# Using the DataExplorer package I will transform the categorical variables
# into dummy variables before fitting any models
german_df <- dummify(german_df)
```

I also wish to create a function which will split up my data into training and testing data randomly into proportions of my choosing. I have created a function below which splits the data into training and test sets with a proportion "x" to be specified by the user. I also set a seed in order to ensure my results are reproducible.

```{r train/test split}

# creating a function to split a data set into training and test sets

train_test <- function(data,x){
  # setting the seed so results can be reproduced
  set.seed(1759)
  
  # selecting x% of rows of n rows randomly
  rand_rows <- sample.int(n = nrow(data),size = round(x*nrow(data),0),replace = FALSE)
  
  # selecting training set
  train <- data[rand_rows,]
  
  # selecting test set
  test <- data[-rand_rows,]
  
  return(list(train,test))
}

# running function and outputting result
result_train_test <- train_test(german_df,0.8)

# creating a training set with 80% of the data
train_german_df <- result_train_test[[1]]

# creating a test set with the remaining 20%
test_german_df <- result_train_test[[2]]
```

## Classification Functions
### Logistic Regression

Below I create the logistic regression function which builds a logistic model using the training data set and then predicts if a customer is "bad" or "good" based on the test set. The function then computes various performance metrics such as accuracy which is the number of true positives (TP) plus true negatives (TN) over the  total number of positive and negative samples (i.e. the total number of rows in the test data). It also computes the mis-classification rate (1 - accuracy), the true and false positive rates and the AUC, (area under the curve) which is one of the most important performance metrics for binary classifiers. The function also computes the TPR and FPR rates for a wide variety of TPR values which we can plot on a ROC curve later on.

```{r logistic regression function, warning=FALSE}
# Defining a function to perform a logistic regression

# Function will take 2 arguments, train and test
# train is the training data set used to train the model
# test is the testing data set used to evaluate model performance
# For the purposes of this project I assume there are no missing values
# in the data set and also assume all categorical variables have been
# converted to binary dummy variables

log_reg_fun <- function(train,test){
  # We can use the glm function along with family = binomial to train a logistic model
  mod <- glm(formula = Credit_Rating ~ .,
             family = binomial(link = "logit"),
             data = train)
  
  # predicting the credit rating on the test data
  res <- predict(mod,newdata = test,type = "response")
  
  # We want to predict whether a customer has a "bad" or "good" credit rating
  res <- ifelse(res > 0.5,1,0)
  
  # computing model accuracy
  mod_acc <- round(mean(res == test$Credit_Rating),3)
  
  # computing misc-classification rates
  mod_mis_class <- round(mean(res != test$Credit_Rating),3)
  
  # creating a confusion matrix
  cm_table <- table("target" = test$Credit_Rating,
                         "prediction" = res)
  row.names(cm_table) <- c("Good","Bad")
  colnames(cm_table) <- c("Good","Bad")
  
  # calculating tp,fn,fp,tn counts (true positive,false negative,false positive,true negative)
  # tp = target = 1, pred = 1
  tp <- cm_table[2,2]
  # fn = target = 1, pred = 0
  fn <- cm_table[2,1]  
  # fp = target = 0, pred = 1
  fp <- cm_table[1,2]
  # tn = target = 0, pred = 0
  tn <- cm_table[1,1]  
  
  # calculating TPR and FPR rates (true and false positive rates)
  # also known as recall and fallout
  TPR <- round(tp / (tp+fn),3)
  FPR <- round(fp / (fp+tn),3)
  
  # calculating auc (area under the curve) using the ROCR package
  
  # predicting the model on the test data again
  pred <- predict(mod,newdata = test,type = "response")
  # creating a prediction object using the ROCR package
  preds <- prediction(pred,test$Credit_Rating)
  # calculating auc
  auc <- performance(preds,measure = "auc")
  auc <- round(auc@y.values[[1]],3)
  
  # calculating x and y values for the ROC plot (plot fo tpr vs fpr for various values of tpr)
  roc <- performance(preds,measure = "tpr",x.measure = "fpr")
  
  # collating and returning the results
  return(list(mod_acc,mod_mis_class,TPR,FPR,auc,roc,cm_table))
}

```

### Decision Tree

Similar to the logistic regression function above, I have produced a function to perform a decision tree classification algorithm and compile the results. I return the same metrics in order to be able to conduct a comparison between the two different methods.

```{r Decision tree function}
# Defining a function to perform a decision tree classification
# Uses the rpart package to do this

# Function will take 2 arguments, train and test
# train is the training data set used to train the model
# test is the testing data set used to evaluate model performance
# For the purposes of this project I assume there are no missing values
# in the data set and also assume all categorical variables have been
# converted to binary dummy variables

tree_fun <- function(train,test){
  # We can use the rpart function along with method = class to train a decision tree model
  mod <- rpart(formula = Credit_Rating ~ .,
               data = train,
               method = "class") # classification problem
  
  # predicting the credit rating on the test data
  res <- predict(mod,newdata = test,type = "class")
  
  # computing model accuracy
  mod_acc <- round(mean(res == test$Credit_Rating),3)
  
  # computing misc-classification rates
  mod_mis_class <- round(mean(res != test$Credit_Rating),3)
  
  # creating a confusion matrix
  cm_table <- table("target" = test$Credit_Rating,
                         "prediction" = res)
  row.names(cm_table) <- c("Good","Bad")
  colnames(cm_table) <- c("Good","Bad")
  
  # calculating tp,fn,fp,tn counts (true positive,false negative,false positive,true negative)
  # tp = target = 1, pred = 1
  tp <- cm_table[2,2]
  # fn = target = 1, pred = 0
  fn <- cm_table[2,1]  
  # fp = target = 0, pred = 1
  fp <- cm_table[1,2]
  # tn = target = 0, pred = 0
  tn <- cm_table[1,1]  
  
  # calculating TPR and FPR rates (true and false positive rates)
  # also known as recall and fallout
  TPR <- round(tp / (tp+fn),3)
  FPR <- round(fp / (fp+tn),3)
  
  # calculating auc (area under the curve) using the ROCR package
  
  # predicting the model on the test data again this time using
  # type = prob in order to get the probabilities of each observation
  # being from the "bad" class (this is in the second column)
  pred <- predict(mod,newdata = test,type = "prob")[,2]
  # creating a prediction object using the ROCR package
  preds <- prediction(pred,test$Credit_Rating)
  # calculating auc
  auc <- performance(preds,measure = "auc")
  auc <- round(auc@y.values[[1]],3)
  
  # calculating x and y values for the ROC plot (plot fo tpr vs fpr for various values of tpr)
  roc <- performance(preds,measure = "tpr",x.measure = "fpr")
  
  # collating and returning the results
  return(list(mod_acc,mod_mis_class,TPR,FPR,auc,roc,cm_table))
}

```

Next I will create the function which will perform both the logistic and decision tree algorithms, collate their results together and finally create an S3 class object which I will write summary methods for. 

```{r overall regression function, warning=FALSE}
# Producing a function to collate the results of the regressions 

classification_summary <- function(train,test){
  
  # Performing Logistic classification
  log_res <- log_reg_fun(train,test)
  
  # Performing Decision Tree classification
  tree_res <- tree_fun(train,test)
  
  # Model Accuracy
  Accuracy <- c("Logistic" = log_res[[1]],
                "Decision Tree" = tree_res[[1]])
  
  # Model Error (mis-classification rates)
  Error <- c("Logistic" = log_res[[2]],
             "Decision Tree" = tree_res[[2]])
  
  # Model TPR
  TPR <- c("Logistic" = log_res[[3]],
             "Decision Tree" = tree_res[[3]])
  
  # Model FPR
  FPR <- c("Logistic" = log_res[[4]],
             "Decision Tree" = tree_res[[4]])
  
  # Model AUC
  AUC <- c("Logistic" = log_res[[5]],
             "Decision Tree" = tree_res[[5]])
  
  # Model ROC Curves
  ROC_Curve <- c("Logistic" = log_res[[6]],
             "Decision Tree" = tree_res[[6]])
  
  # Model Confusion Matrices
  Confusion_Matrix <- list(log_res[[7]],tree_res[[7]])
  
  # Summary table of results
  Summary_Table <- rbind(Accuracy,Error,TPR,FPR,AUC)
  
  # generating results
  results <- list("Accuracy" = Accuracy,
                  "Mis-classification Error" = Error,
                  "TPR" = TPR,
                  "FPR" = FPR,
                  "AUC" = AUC,
                  "ROC Curve points" = ROC_Curve,
                  "Confusion Matrices" = Confusion_Matrix,
                  "Summary Table" = Summary_Table)
  
  # setting the class of results to be class = "classification"
  class(results) <- "classification"
  
  invisible(results)
} 

# setting a variable equal to the output of the function to demonstrate
# the summary methods later on
reg_results <- classification_summary(train_german_df,test_german_df)
x <- classification_summary(train_german_df,test_german_df)

# printing the class of the reg_results variable - "classification"
print(class(reg_results))

```

## Summary Methods for S3 Class
### Print Method  

I will now write three functions to produce print, summary and plot methods for my newly developed "classification" class. The first of these functions, print, will be fairly basic and simply output the accuracy of the logistic regression model and decision tree model. I then demonstrate the use of this print function on the "reg_results" variable which was set equal to the output of my main summary function and has a class of "classification".

```{r print method}
# Creating a print method for the "classification" class
print.classification <- function(x){
  cat("The accuracy of the classification models are shown below.\n\n")
  print(x$Accuracy)
  cat("\n")
  if (x$Accuracy["Logistic"] > x$Accuracy["Decision Tree"]) {
    cat("The Logistic Model is more accurate")
  } else {
    cat("The Decision Tree Model is more accurate")
  }
}

# Demonstrating the print function on the summary results variable
print(reg_results)
```

As we can see from the output of the print method above, the logistic model has an accuracy of 75% whereas the decision tree has an accuracy of 73.5%. In terms of the logistic model an accuracy of 75% means that model predicted 3 out of every 4 of the customer's credit ratings correctly and mis-classified 1 out of every 4. The decision tree model is slightly worse but still relatively good. An accuracy of 75% is not bad however I believe model accuracy could be improved if K-fold cross validation was performed during the training phase in order to fine-tune the parameters of the model.   

### Summary Method    

I will now create the summary method for my the "classification" S3 class. I want the summary method to display various performance metrics such as accuracy, mis-classification rate, the true and false positive rates which are the number of correctly identified positives divided by the total number of positive samples. In the case of the German Credit data, this would be the total number of correctly identified "bad" credit ratings divided by the total number of "bad" credit ratings in the test data set. The false positive rate follows the same logic, the number of incorrectly identified "good" credit ratings divided by the total number of "good" credit ratings in the test data. I also wish to output the area under the curve or (AUC).  

The AUC is a very powerful metric in terms of binary classifier performance and represents the area underneath the ROC (receiver operating characteristic curve). The AUC can take values on [0,1] and can be interpreted as the expected proportion of positives i.e. "bad" credit customers ranked before a uniformly drawn random negative i.e. "good" credit customer. A model which randomly assigns customers as either "bad" or "good" would have an AUC of 0.5 and a model which perfectly predicts customer's credit ratings would have an AUC of 1. The larger the AUC the better as this means are model can better predict whether or not a customer has a "bad" credit rating.

```{r summary method}
# Creating a summary method for the "classification" class
summary.classification <- function(x){
  cat("The summary performance metrics of the models are shown below.\n\n")
  print(x$`Summary Table`)
  cat("\n")
}

# Demonstrating the summary function on the reg_results variable
summary(reg_results)
```

As we can see above the logistic regression outperforms the decision tree in every metric aside from the true positive rate (TPR). This means the decision tree model is actually better at predicting "bad" credit rating customer's than the logistic regression model and this is the more important classification due to its commercial nature. What I mean by this is that it is less of a problem to classify a "good" customer as being "bad" than classifying a "bad" customer as being "good". The reason for this is that we will likely lend money out to the "good" customers if we are a bank and if we have incorrectly identified a "bad" customer as being "good" then we will lend out money to a customer with a bad credit rating and will be far more likely to lose money on that loan. Conversely, if we identified a "good" customer as being "bad" we will just not lend any money to this customer. While there is an opportunity cost in terms of interest we could earn by loaning money to a good customer, there is no actual credit risk to the bank as we haven't lent any money.

We also see that the AUC's of both the logistic and decision tree is above 0.7 and suggests that both models perform relatively well at classifying customer's credit ratings. As the logistic model has a higher AUC and accuracy we can say that this model overall, is a better classifier than the decision tree model. It must be noted that both of these models have been fitted without any model tuning and as such we could likely achieve better results if we first tune the models (using tools such as cross-validation) in order to optimise model parameters.

### Plot Method  

Finally, I will produce a plot method for the "classification" class. I want the plot function to plot the confusion matrices for each of the models and also the ROC curves. The confusion matrix is a table of predicted vs actual counts for each of the credit ratings "bad" and "good". The ROC curve is a plot of the TPR (true positive rate) against the FPR (false positive rate) for various values of the TPR.

```{r plot method}
plot.classification <- function(x){
  # firstly plot the confusion matrices
  cm_logistic <- conf_mat(data = x$`Confusion Matrices`[[1]])
  # specifying "heatmap" type and also adding titles and colour scheme
  cm_logistic_plot <- autoplot(cm_logistic,type = "heatmap") +
    ggtitle("Logistic Regression") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_gradient2(low = "blue",high = "red")
  
  cm_tree <- conf_mat(data = x$`Confusion Matrices`[[2]])
  # specifying "heatmap" type and also adding titles and colour scheme
  cm_tree_plot <- autoplot(cm_tree,type = "heatmap") +
    ggtitle("Decision Tree") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_gradient2(low = "blue",high = "red")
  
  
  
  # Next plot the ROC curves already computed by the ROCR package
  par(mfrow = c(1,2))
  plot(plot_grid(cm_logistic_plot,cm_tree_plot))
  
  par(mfrow = c(1,2))
  plot(x$`ROC Curve points`[[1]],main = "Logistic - ROC Curve",colorize = TRUE)
  plot(x$`ROC Curve points`[[2]],main = "Decision Tree - ROC Curve",colorize = TRUE)
     
}

# Demonstrating the use of the plot function for the classification class
plot(reg_results)
```

Finally, we can see the output of the plot method above. The plot method outputs each of the model's confusion matrices which is a table of the predicted customer ratings vs the actual (target) customer ratings. We can see that the logistic model correctedly predicted 125 "goos" customers compared to the 117 for the decision tree model. This is related to the true positive rates (TPRs) we observed using the summary method. As we noted previously, the decision tree is better at predicting "bad" customers, and managed to correctly predict 30 bad customer's credit ratings compared to only 25 for the logistic model.

The plot method also produces the ROC (receiver operating characteristic) curves for the two regression models. The ROC curve is a plot of the FPR vs the TPR for various threshold values (i.e. various probabilities at which we classify a "bad" customer). The ROC is the curve for which the area under the curve (AUC) is calculated. The higher the AUC value i.e. the more to the top left of the graph the ROC curve goes, the better the model is at distinguishing between classes. It is difficult to see which model is actually performing better based on the plots above, and that's why it is a good idea to export the AUC as well as the plot as we did in the summary method. 

# Appendices

## Bibliography
In this section I cite and provide a reason for using each of the packages that were not dealt with extensively during the course.

### Scales
I use the scales package throughout the project (particularly in part 1) to adjust the x and y axis values and in particular to convert them to percentages for some of the graphs I produced in part 1.  

```{r scales citation}
citation("scales")
```
### Cowplot
I use the cowplot package to arrange several ggplot2 plots in the same figure. The function in particular I used throughout the project was the plot_grid function.

```{r Cowplot citation}
citation("cowplot")
```

### KableExtra
I use the KableExtra in some of the early tables in part 1 in order to fit them nicely on the PDF with pleasant formatting.

```{r KableExtra citation}
citation("kableExtra")
```

### ROCR
I use the ROCR package in part 3 in order to calculate the AUC (area under the curve statistic) and also to plot the ROC (Receiver Operating Characteristic) curves.

```{r ROCR citation}
citation("ROCR")
```

### rpart
I use the rpart package in order to fit a decision tree model in part 3.

```{r rpart citation}
citation("rpart")
```

### yardstick
I use the yardstick package in order to graph my basic homemade confusion matrices and make them look more professional.

```{r yardstick citation}
citation("yardstick")
```

